---
title: "ANES 2020 Derived Variable Codebook"
output:
  html_document:
     css: "style.css"
     toc: true
     toc_depth: 4
     toc_float: true
     self_contained: true
  github_document:
     toc: true
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(janitor)
library(kableExtra)
library(knitr)

anes_2020 <- readRDS(here::here("Data", "anes_2020.rds")) %>% 
   haven::zap_labels()
```

The full codebook with the original variables is available on <a href="https://electionstudies.org/wp-content/uploads/2022/02/anes_timeseries_2020_userguidecodebook_20220210.pdf" target="_blank">electionstudies.org</a>.

## Weighting variables

### V200010b

Full sample post-election weight.

### V200010d

Full sample variance stratum.

### V200010c

Full sample variance unit.

## InterviewMode

MODE OF INTERVIEW: PRE-ELECTION INTERVIEW

```{r variable-InterviewMode}
anes_2020 %>%
   count(V200002, InterviewMode) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   janitor::adorn_totals() %>%
   kbl() %>%
   kable_minimal()
```
Variables used: V200002

## Age

RESPONDENT AGE

```{r variable-Age}
anes_2020 %>%
   count(AgeGroup) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201507x

## Gender

PRE: WHAT IS YOUR (R) SEX? [REVISED]

What is your sex?

```{r variable-Gender}
anes_2020 %>%
   count(V201600, Gender) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201600

## RaceEth

PRE: SUMMARY: R SELF-IDENTIFIED RACE/ETHNICITY

```{r variable-RaceEth}
anes_2020 %>%
   count(V201549x, RaceEth) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201549x

## PartyID

PRE: SUMMARY: PARTY ID

```{r variable-PartyID}
anes_2020 %>%
   count(V201231x, PartyID) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201231x

## Education

```{r variable-Education}
anes_2020 %>%
   count(V201510, Education) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201510

## Income

PRE: SUMMARY: TOTAL (FAMILY) INCOME

```{r variable-Income}
anes_2020 %>%
   count(V201617x, Income) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

```{r variable-Income7}
anes_2020 %>%
   count(V201617x, Income7) %>% 
   group_by(Income7) %>% 
   summarise(n = sum(n)) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201617x

## CampaignInterest

PRE: HOW INTERESTED IN FOLLOWING CAMPAIGNS

Some people don’t pay much attention to political campaigns. How about you? Would you say that you have been very much interested, somewhat interested or not much interested in the political campaigns so far this year?

```{r variable-CampaignInterest}
anes_2020 %>% 
   count(V201006, CampaignInterest) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201006

## TrustGovernment

PRE: HOW OFTEN TRUST GOVERNMENT IN WASHINGTON TO DO WHAT IS RIGHT [REVISED]

How often can you trust the federal government in Washington to do what is right?

```{r variable-TrustGovernment}
anes_2020 %>%
   count(V201233, TrustGovernment) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201233

## TrustPeople

PRE: HOW OFTEN CAN PEOPLE BE TRUSTED

Generally speaking, how often can you trust other people?

```{r variable-TrustPeople}
anes_2020 %>%
   count(V201237, TrustPeople) %>% 
   rename(Label = V201237) %>%
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201237

## VotedPres2016

PRE: DID R VOTE FOR PRESIDENT IN 2016

Four years ago, in 2016, Hillary Clinton ran on the Democratic ticket against Donald Trump for the Republicans. Do you remember for sure whether or not you voted in that election?

*Revised:* Four years ago, in 2016, Hillary Clinton ran on the Democratic ticket against Donald Trump for the Republicans. We talk to many people who tell us they did not vote. And we talk to a few people who tell us they did vote, who really did not. We can tell they did not vote by checking with official government records. What about you? If we check the official government voter records, will they show that you voted in the 2016 presidential election, or that you did not vote in that election?

```{r variable-VotedPres2016}
anes_2020 %>%
   count(V201101, V201102, VotedPres2016) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201101, V201102

## VotedPres2016_selection

PRE: RECALL OF LAST (2016) PRESIDENTIAL VOTE CHOICE

Which one did you vote for?

```{r variable-VotedPres2016_selection}
anes_2020 %>%
   count(V201103, VotedPres2016_selection) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V201103

## VotedPres2020

PRE-POST: SUMMARY: VOTER TURNOUT IN 2020

```{r variable-VotedPres2020}
anes_2020 %>%
   count(V202109x, VotedPres2020) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

Variables used: V202109x

## VotedPres2020_selection

POST: FOR WHOM DID R VOTE FOR PRESIDENT

```{r variable-VotedPres2020_selection}
anes_2020 %>%
   count(V202073, VotedPres2020_selection) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```

## EarlyVote2020 / VotedPres2020

PRE: SUMMARY: REGISTRATION AND EARLY VOTE STATUS

```{r variable-EarlyVote2020}
anes_2020 %>%
   count(V201025x, VotedPres2020, EarlyVote2020) %>% 
   mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
   kbl() %>%
   kable_minimal()
```
