---
title: "Tidy Survey Analysis in R using the srvyr Package"
subtitle: "Short Course Day 1 - Categorical Data"
author:
   - Stephanie Zimmer, Abt Associates
   - Rebecca Powell, RTI International
   - Isabella Velásquez, RStudio
date: "TBD 2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, tidy = FALSE)
```



```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1E4F96",
  secondary_color = "#00A3E0",
  inverse_header_color = "#FFFFFF"
)
```

class: inverse center middle
# Introduction

---


```{css, echo = FALSE}
.small .remark-code { /*Change made here*/
  font-size: 80% !important;
}
.smaller .remark-code { /*Change made here*/
  font-size: 70% !important;
}
```

## Overview

- At the end of this course, you should be able to 
  - Calculate point estimates and their standard errors with survey data 
      - Proportions, totals, and counts
      - Means, quantiles, and ratios
  - Perform t-tests and chi-squared tests
  - Fit regression models
  - Specify a survey design in R to create a survey object
  
- We will not be going over the following but provide some resources at the end
  - Weighting (calibration, post-stratification, raking, etc.)
  - Survival analysis
  - Nonlinear models



---
## Overview: Course Roadmap

- Get familiar with RStudio Cloud with a warm-up exercise using the tidyverse (today)

- Introduce the survey data we'll be using in the course (today)

- Analysis of categorical data with time for practice (today)

- Analysis of continuous data with time for practice (day 2)

- Specify a survey design object in R with exercises (day 3)

- Resources for other survey analysis topics (day 3)


---
## Logistics

- We will be using RStudio Cloud today to ensure everyone has access

    - Sign-up for a free RStudio Cloud account 
    - Access the project and files via link in email and Zoom chat
    - Click "START" to open the project and get started
    - Rstudio Cloud has the same features and appearance as RStudio for ease of use

- All slides and code are available on GitHub: https://github.com/tidy-survey-r/tidy-survey-short-course

???
Github repo is for future reference, all material on RStudio cloud

---
## Intro to RStudio Cloud: Penguins!!

- Using `palmerpenguins` data for warm-up exercises

- Data were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network.

- Access data through `palmerpenguins` package https://github.com/allisonhorst/palmerpenguins/


####If you are using your own RStudio environment:
- Make sure you have `tidyverse`, `here`, and `palmerpenguins` installed

```{r inst_packages, error=FALSE, warning=FALSE, eval=FALSE}
# Run package installation if you don't have these packages already
# As a reminder, installing takes package from internet to your computer 
# and only needs to be done once, not each session

install.packages(c("tidyverse", "here", "palmerpenguins"))
```

---
## Intro to RStudio Cloud: Penguins!!

- Load `tidyverse`, `here`, and `palmerpenguins`

- Look at the penguins dataset using `glimpse`

```{r load_pack1, error=FALSE, warning=FALSE}
library(tidyverse) # for tidyverse
library(here) # for file paths
library(palmerpenguins) # for warm-up data
glimpse(penguins)
```

---
## Warm-up Exercises: WarmUpExercises.Rmd

- <b>Let's open RStudio cloud and do some warm-up examples</b>

- Explore the penguins data
  - How many penguins of each species are there? <i>We will do this one together</i>
  - How many penguins of each species and sex are there? Hint: use `count`


---
## Ex. 1: How many penguins of each species are there? 

```{r peng1}
penguins %>%
   count(species)
```

???
- Using `count` we see there are 3 different species and the number of penguins for each species
---
## Ex. 2: How many penguins of each species and sex are there?

```{r peng2}
penguins %>%
   count(species, sex)
```
???
- `count` can take more than one variable to get a cross-tabs between the two variables

---
class: inverse center middle
# Survey Datasets
---
## American National Election Studies (ANES) 2016 

- Pre and post election surveys

- Fielded almost every 2 years since 1948

- Topics include voter registration status, candidate preference, opinions on country and government, party and ideology affiliation, opinions on policy, news sources, and more

- Collaboration of Stanford, University of Michigan - funding by the National Science Foundation

- <b>Target Population</b>: US citizens, 18 and older living in US 

- <b>Mode</b>: FTF with CASI and Web

- <b>Sample Information</b>: Pseudo-strata and pseudo-cluster included for variance estimation

https://electionstudies.org/

???
- We have subset the columns of this data and created derived variables, code in repository
- Chose not to use 2020 data because it is still preliminary

---
class: inverse center middle
# Categorical data analysis
---
## Overview of Survey Analysis using `srvyr` Package

1. Create a `tbl_svy` object using: `as_survey_design` or `as_survey_rep`

2. Subset data (if needed) using `filter` (subpopulations)

3. Specify domains of analysis using `group_by` 

4. Within `summarize`, specify variables to calculate including means, totals, proportions, quantiles and more


<b>Note: We will be teaching this in the reverse order!!!</b>
---
## Weighted Analysis for Categorical Variable

- Functions to use within `summarize` after `group_by`
   - survey_mean
   - survey_total

- Functions to get counts
  - survey_count

???

- we use the same mean and total functions as with continuous variables
- `survey_count` is new
    - has a similar structure as the standard (non-survey) version of count


---
## Set-up for Analysis
- `srvyr` package uses tidy-syntax but uses the `survey` package behind it to do calculations

- If using your own RStudio environment, install both packages:
```{r inst_srv, eval=FALSE}
# Install survey and srvyr packages

remotes::install_github("bschneidr/survey", ref = "c217689")
install.packages("srvyr")
```


- First, we will set-up a design object and later talk about what it means
```{r anes_des, error=FALSE, warning=FALSE}
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

anes <- read_rds(here("Data", "anes.rds")) %>%
   mutate(Weight=Weight/sum(Weight)*224059005) 
# adjust weight to sum to citizen pop, 18+ in Nov 2016 per ANES methodology documentation
anes_des <- anes %>%
   as_survey_design(weights = Weight,
                    strata = Stratum,
                    ids = VarUnit,
                    nest = TRUE)

```
???
- need to install github version of survey package if you want CIs with quantiles
- American National Election Studies
- provides weights that sum to the sample, but we want to get population estimates
    - need to adjust the weight to get it to the population count
- we will cover setting up the sample design object later

---
## `survey_count` Syntax

- `survey_count` functions similarly to `count` in that it is <b>NOT</b> called within `summarize`

- Produces weighted counts and variance of your choice of those counts
```{r survey_count_syn, eval=FALSE, tidy=FALSE}
survey_count(
  x,
  ...,
  wt = NULL,
  sort = FALSE,
  name = "n",
  .drop = dplyr::group_by_drop_default(x),
  vartype = c("se", "ci", "var", "cv")
)
```
???
- similar to count in that it takes one or many variables
- can change the variance type

---
## `survey_count` Example

- Cross-tab of population in each age group and gender
```{r survey_count_ex}
anes_des %>%
  survey_count(AgeGroup, name="N")

```
???
- `survey_count` is placed on its own like `count`
- it does NOT go in a `summarize` function
- can take multiple variables
- can change the output count name, `n` is the default

---
## `survey_mean` and `survey_total` within `summarize`
- Specify the sample design,
- then specify the crosstab in `group_by`,
- then `survey_mean` used with no x (variable) calculates a proportion of groups within `summarize`, or
- `survey_total` used with no x (variable) calculates a population count estimate within `summarize`

---
## `survey_mean` and `survey_total` Examples

Looking at population by age group as done with `survey_count`.
```{r survey_p_ex1}
anes_des %>%
  group_by(AgeGroup) %>%
  summarize(
    p=survey_mean(),
    N=survey_total(),
    n=unweighted(n()), # this gets unweighted counts aka sample sizes
    .groups="drop" # summarize option to remove groups
  )
```
???
- to get proportions we use `group_by` and `survey_mean`
- also use `survey_total` to get a population count estimate as before

---
## Conditional proportions with more than one group

- Specifying more than one group calculates conditional proportions
- Example: people voting in 2012 and 2016

```{r survey_p_cond, tidy=FALSE}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(VotedPres2012, VotedPres2016) %>%
  summarize(
    p=survey_mean(),
    N=survey_total(),
    n=unweighted(n()), 
    .groups="drop"
  )
```
???
- Note that this is the proportion of voting in 2016 by whether people voted in 2012
- What if we don't want conditional proportions?

---
## Joint proportions with more than one group

- Specify an interaction to get joint distribution - use `interact` within `group_by`
- Example: people voting in 2012 and 2016

```{r survey_p_joint}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(interact(VotedPres2016, VotedPres2012)) %>% #<<
  summarize(
    p=survey_mean(),
    N=survey_total(),
    .groups="drop"
  )
```
???
- We add an interaction for the groups
- This outputs the joint distribution

---
## Proportions with Design Effects

```{r survey_p_deff}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(VotedPres2012, VotedPres2016) %>%
  summarize(
    p=survey_mean(deff=TRUE),
    N=survey_total()
  )
```
???
- Use `deff=TRUE` option in the `survey_mean` function

---
## Practice on your own

- Open CategoricalExercises.Rmd and work through Part 1

- We will take 7 minutes. Use this time for the exercises