---
title: "Tidy Survey Analysis in R using the srvyr Package"
subtitle: "AAPOR 2021 Short Course"
author:
   - Stephanie Zimmer, RTI International
   - Rebecca Powell, RTI International
date: "2021-05-06"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, tidy = FALSE)
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1E4F96",
  secondary_color = "#00A3E0",
  inverse_header_color = "#FFFFFF"
)
```

class: inverse center middle
# Introduction

---

```{css, echo = FALSE}
.small .remark-code { /*Change made here*/
  font-size: 80% !important;
}
.smaller .remark-code { /*Change made here*/
  font-size: 70% !important;
}
```

## Overview

- At the end of this course, you should be able to 
  - Calculate point estimates and their standard errors with survey data 
      - Means & Proportions
      - Totals
      - Quantiles
  - Perform t-tests and chi-squared tests
  - Fit regression models
  - Specify a survey design in R to create a survey object
  
- We will not be going over the following but provide some resources at the end
  - Weighting (calibration, post-stratification, raking, etc.)
  - Survival analysis
  - Nonlinear models



---
## Overview: Course Roadmap

- Get familiar with RStudio Cloud with a warm-up exercise using the tidyverse

- Introduce the survey data we'll be using in the course

- Analysis of continuous data with time for practice

- Analysis of categorical data with time for practice

- Specify a survey design object in R with exercises

- Resources for other survey analysis topics

- Closing

---
## Logistics

- We will be using RStudio Cloud today to ensure everyone has access

    - Sign-up for a free RStudio Cloud account 
    - Access the project and files via link in email and Zoom chat
    - Click "START" to open the project and get started
    - Rstudio Cloud has the same features and appearance as RStudio for ease of use

- All slides and code are available on GitHub: https://github.com/szimmer/tidy-survey-aapor-2021

???
Github repo is for future reference, all material on RStudio cloud
---
## Intro to RStudio Cloud: Penguins!!

- Using `palmerpenguins` data for warm-up exercises

- Data were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network.

- Access data through `palmerpenguins` package https://github.com/allisonhorst/palmerpenguins/


####If you are using your own RStudio environment:
- Make sure you have `tidyverse`, `here`, and `palmerpenguins` installed

```{r inst_packages, error=FALSE, warning=FALSE, eval=FALSE}
# Run package installation if you don't have these packages already
# As a reminder, installing takes package from internet to your computer 
# and only needs to be done once, not each session

install.packages(c("tidyverse", "here", "palmerpenguins"))
```

---
## Intro to RStudio Cloud: Penguins!!

- Load `tidyverse`, `here`, and `palmerpenguins`

- Look at the penguins dataset using `glimpse`

```{r load_pack1, error=FALSE, warning=FALSE}
library(tidyverse) # for tidyverse
library(here) # for file paths
library(palmerpenguins) # for warm-up data
glimpse(penguins)
```

---
## Warm-up Exercises: WarmUpExercises.Rmd

- <b>Let's open RStudio cloud and do some warm-up examples</b>
  - We will do one together and then give you 5-minutes to work through other examples and get familiar with RStudio Cloud

- Explore the penguins data
  - How many penguins of each species are there? <i>We will do this one together</i>
  - How many penguins of each species and sex are there? Hint: use `count`
  - What is the mean length of flipper by species? Hint: use `group_by` and `summarize`
  - What is the mean flipper length by species and sex? 

- More advanced warm-up
  - Fit a simple linear regression between body mass and flipper length.
  - Test whether the average flipper length is significantly different between male and female penguins. Use t-test, lm, or glm

---
## Ex. 1: How many penguins of each species are there? 

```{r peng1}
penguins %>%
   count(species)
```

???
- Using `count` we see there are 3 different species and the number of penguins for each species
---
## Ex. 2: How many penguins of each species and sex are there?

```{r peng2}
penguins %>%
   count(species, sex)
```
???
- `count` can take more than one variable to get a cross-tabs between the two variables

---
## Ex. 3: What is the mean length of flipper by species? 

```{r peng3}
penguins %>%
   group_by(species) %>%
   summarize(
     MeanFlipperLength=mean(flipper_length_mm, 
                            na.rm=TRUE))
```
???
- `group_by` allows us to look at metrics by different subgroups like species
- when using `group_by` follow it with `summarize` to get metrics (like average) at the group level
- `na.rm=TRUE` removes missing data from the calculation
    - forgetting this argument will result in a value of `NA` as the function will try to average missing data

---
## Ex. 4: What is the mean flipper length by species and sex?

```{r peng4}
penguins %>%
   group_by(species, sex) %>%
   summarize(
     MeanFlipperLength=mean(flipper_length_mm,
                            na.rm=TRUE))
```
???
- As with `count`, you can `group_by` multiple variables

---
## Advanced Ex. 1: Linear regression (body mass & flipper length)
.small[
```{r pengad1}
mod1 <- lm(body_mass_g ~ flipper_length_mm, data=penguins)
summary(mod1)
```
]
???
- use `lm` (linear model) function
- equation is written as y-variable ~ x-variables

---
## Advanced Ex. 2: Flipper length differences by sex: t-test

```{r pengad2a}
t.test(flipper_length_mm ~ sex, data=penguins)
```
???
- ~ also used in `t.test` to separate what we want to measure (our y) and the groups of interest (our x)
---
## Advanced Ex. 2: Flipper length differences by sex: lm 
.small[
```{r pengad2b}
mod3 <- lm(flipper_length_mm ~ sex, data=penguins)
summary(mod3)
```
]
---
## Advanced Ex. 2: Flipper length differences by sex: glm 
.small[
```{r pengad2c}
mod4 <- glm(flipper_length_mm ~ sex, data=penguins)
summary(mod4)
```
]
???
- `glm` takes the same arguments as `lm`, but is more flexible for working with non-normal data
---
class: inverse center middle
# Survey Datasets
---
## Residential Energy Consumption Survey (RECS) 2015

- Energy consumption/expenditures collected through energy suppliers

- Fielded 14 times between 1950 and 2015

- Topics include appliances, electronics, heating, a/c, temperatures, water heating, lighting, energy bills, respondent demographics, and energy assistance

- Funded by the Energy Information Administration

- <b>Target Population</b>: Primary occupied housing units in the US

- <b>Mode</b>: In-person, paper, and web interview mode

- <b>Sample Information</b>: BRR Replicate weights included for variance estimation

https://www.eia.gov/consumption/residential/index.php

???
- We have subset the columns of this data and created derived variables, code in repository
---
## American National Election Studies (ANES) 2016 

- Pre and post election surveys

- Fielded almost every 2 years since 1948

- Topics include voter registration status, candidate preference, opinions on country and government, party and ideology affiliation, opinions on policy, news sources, and more

- Collaboration of Stanford, University of Michigan - funding by the National Science Foundation

- <b>Target Population</b>: US citizens, 18 and older living in US 

- <b>Mode</b>: FTF with CASI and Web

- <b>Sample Information</b>: Pseudo-strata and pseudo-cluster included for variance estimation

https://electionstudies.org/

???
Chose not to use 2020 data because it is still preliminary

---
class: inverse center middle
# Continuous data analysis
---
## Overview of Survey Analysis using `srvyr` Package

1. Create a `tbl_svy` object using: `as_survey_design` or `as_survey_rep`

2. Subset data (if needed) using `filter` (subpopulations)

3. Specify domains of analysis using `group_by` 

4. Within `summarize`, specify variables to calculate including means, totals, proportions, quantiles and more


<b>Note: We will be teaching this in the reverse order!!!</b>
---
## Set-up for Analysis
- `srvyr` package uses tidy-syntax but uses the `survey` package behind it to do calculations

- If using your own RStudio environment, install both packages:
```{r inst_srv, eval=FALSE}
# Install survey and srvyr packages

remotes::install_github("bschneidr/survey", ref = "c217689")
install.packages("srvyr")
```

- First, we will set-up a design object and later talk about what it means
```{r recs_des, error=FALSE, warning=FALSE}
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

recs <- read_rds(here("Data", "recs.rds"))

recs_des <- recs %>%
   as_survey_rep(weights=NWEIGHT,
                 repweights=starts_with("BRRWT"),
                 type="Fay",
                 rho=0.5,
                 mse=TRUE)

```
???
- need to install github version of survey package if you want CIs with quantiles
---
## Weighted Analysis for Continuous Variables

- Common functions for continuous summaries
   - survey_mean
   - survey_total (like sum)
   - survey_median
   - survey_quantile
   - survey_ratio

- Always call within `summarize`/`summarise`
---
## `survey_mean` Syntax

```{r survey_mean_syn, eval=FALSE}
survey_mean(
  x,
  na.rm = FALSE,
  vartype = c("se", "ci", "var", "cv"),
  level = 0.95,
  proportion = FALSE,
  deff = FALSE,
  df = NULL,
  ...
)
```

To calculate a survey mean, we use this in `summarize`/`summarise`
```{r survey_mean_syn2, eval=FALSE}
survey_design_object %>%
   summarize(
      mean_varname=survey_mean(x = continuous_varname)
      )
```

???
Only required argument is the variable

---
## `survey_mean` Example 1: Mean dollars spent on energy

This is an example using the `recs_des` survey design object and `survey_mean` function defaults

```{r survey_mean_ex1}
recs_des %>%
   summarize(
      TD_mean=survey_mean(x = TOTALDOL)
      )
```
---
## `survey_mean` Example 2: Mean temperature setting for summer during the day

Run this code. What happens? Why?

```{r survey_mean_ex2, eval=FALSE}
recs_des %>%
   summarize(
      TD_mean=survey_mean(x = SummerTempDay)
      )
```
---
## `survey_mean` Example 2: Mean temperature setting for summer during the day

Run this code. What happens? Why?

```{r survey_mean_ex2_r, error=TRUE}
recs_des %>%
   summarize(
      TD_mean=survey_mean(x = SummerTempDay)
      )
```

<b>How do we fix this code?</b>

???
- missing data in temperature, need `na.rm=TRUE`
---
## `survey_mean` Example 2: Missing data solution

```{r survey_mean_ex2_sol, error=TRUE, tidy=FALSE}
recs_des %>%
   summarize(
      TD_mean = survey_mean(
        x = SummerTempDay, 
        na.rm = TRUE )#<<
      )
```

---
## `survey_median` Syntax

```{r survey_median_syn, eval=FALSE}
survey_median(
  x,
  na.rm = FALSE,
  vartype = c("se", "ci"),
  level = 0.95,
  df = NULL,
  ...
)
```

???
Only required argument is the variable


---
## `survey_median` Example: Median temperature setting for summer during day

Fill in the blank:

```{r survey_median_fib, eval=FALSE}
recs_des %>%
   summarize(
      TD_median=survey_median(x=_________,
                          na.rm=_________)
      )
```

--

```{r survey_median_fib_sol}
recs_des %>%
   summarize(
      TD_median=survey_median(x=SummerTempDay,
                              na.rm=TRUE)
      )
```



---
## `survey_quantile` Syntax

```{r survey_quantile_syn, eval=FALSE}
survey_quantile(
  x,
  quantiles, #<<
  na.rm = FALSE,
  vartype = c("se", "ci", "var", "cv"),
  level = 0.95,
  df = NULL,
  ...
)
```

???
- need both the variable and the quantiles in a vector e.g. (c(.25, .75))
---
## `survey_quantile` Example 1: 1st and 3rd quantile of dollars spent on energy

```{r survey_quantile_ex1, error=TRUE}
recs_des %>%
   summarize(
      Spent=survey_quantile(
        x = TOTALDOL,
        quantiles = c(.25, .75)) #<<
      )
```
???
- This estimates the 25th and 75th percentile

---
## `survey_quantile` Example 2: 1st and 3rd quantile of dollars spent on energy now with confidence interval

```{r survey_quantile_ex2, error=TRUE}
recs_des %>%
   summarize(
      Spent=survey_quantile(x = TOTALDOL,
                            quantiles = c(.25, .75),
                            vartype = "ci" #<<
         )
      )
```
---
## `survey_ratio` Syntax

- Note this estimates: $\sum x_i/\sum y_i$ not $\sum \frac{x_i}{y_i}$

```{r survey_ratio_syn, eval=FALSE}
survey_ratio(
  numerator, #<<
  denominator, #<<
  na.rm = FALSE,
  vartype = c("se", "ci", "var", "cv"),
  level = 0.95,
  deff = FALSE,
  df = NULL,
  ...
)
```


---
## `survey_ratio` Example: mean dollars per BTU spent on energy

```{r survey_ratio_ex}
recs_des %>%
   summarize(
      DolPerBTU=survey_ratio(
         numerator = TOTALDOL, #<<
         denominator = TOTALBTU, #<<
         na.rm = TRUE
         )
      )
```
---
## Practice on your own

- Open ContinuousExercises.Rmd and work through Part 1

- We will take 15 minutes. Use this time for the exercises and a break
---
## Weighted Analysis for Continuous Variables: Domain Analysis

- If we want to get estimates by another variable, we need to add a `group_by` statement before doing the analysis.

- Example: Average dollars spent on electricity by whether AC is used

```{r domain_ex}
recs_des %>%
  group_by(ACUsed) %>% #<<
  summarize(
    ElBill=survey_mean(DOLLAREL, 
                       na.rm=TRUE)
  )
```
---
## Domain Analysis: Totals

- If we want the overall electric bill too, use the `cascade` function instead of `summarize`

```{r domain_ex_casc}
recs_des %>%
   group_by(ACUsed) %>%
   cascade(
      ElBill=survey_mean(DOLLAREL, 
                         na.rm=TRUE)
   )

```

???
- Note the overall appears as NA

---
## Domain Analysis: Totals

- Also can add sample and pop sizes

```{r domain_tot}
recs_des %>%
   group_by(ACUsed) %>%
   cascade(
      ElBill=survey_mean(DOLLAREL, na.rm=TRUE),
      N=survey_total(!is.na(DOLLAREL)), #<<
      n=unweighted(sum(!is.na(DOLLAREL))) #<<
   )

```

???
- survey_total gets a weighted total
- unweighted does just that, an unweighted estimate, can also get an unweighted mean or any other stat

---
## Weighted Analysis for Specific Subpopulations

- filtering (subsetting) the data should be done AFTER specifying the design to ensure accurate standard errors

- Use the `filter` function after creating the survey design object and before summarizing

Wrong way:
```{r filter_bad, eval = FALSE}
data %>%
  filter(state=="NC") %>% #<<
  as_survey_design(...) %>%
  summarize(AvgAge=mean(Age))
```

Right way:
```{r filter_good, eval=FALSE}
data %>%
  as_survey_design(...) %>%
  filter(state=="NC") %>% #<<
  summarize(AvgAge=mean(Age))
```

???
- The difference in these two methods occurs when the subpopulation doesn't occur in all strata or PSUs

---
## Subpopulation Example 1: Average electric cost of single family homes

```{r subpop1}
recs_des %>%
  filter(HousingUnitType %in% c("Single-family detached",
                                "Single-family attached")) %>%
  summarize(
    ElBill=survey_mean(DOLLAREL, 
                       na.rm=TRUE)
  )
```

---
## Comparisons with t-tests: `svyttest` Syntax

- t-tests are done in the package `survey` not `srvyr` but you can use the same design object

```{r ttest_syn, eval=FALSE}
svyttest(formula, # outcome~group for two-sample, outcome~0 for one-sample
         design,
         na.rm = FALSE
         ....)
```

???
- Uses standard R formula notation
- will go over examples of 1-sample, 2-sample, and paired t-test

---
## `svyttest` Example 1: One-sample t-test

- I keep my house at 68 degrees at night during the summer. Is this different from the national average?

```{r ttest_ex1}
recs_des %>%
   svyttest(design=.,
            formula=I(SummerTempNight-68)~0,
            na.rm=TRUE)
```

???
- Note the I notation, this does the arithmetic before modeling

---
## `svyttest` Example 2: Comparing two variables

- Do people keep their house the same temperature at night during the summer and the winter?

```{r ttest_ex2}
recs_des %>%
   svyttest(design=.,
            formula=I(SummerTempNight-WinterTempNight)~0,
            na.rm=TRUE)
```

???
- this is a paired t-test
- testing whether the difference is 0 for each household
---
## `svyttest` Example 3: Two-sample t-test

- Are electric bills different between those with and without A/C?

```{r ttest_ex3}
recs_des %>%
   svyttest(design=.,
            formula=DOLLAREL~ACUsed,
            na.rm=TRUE)
```



---
## Linear Regression or ANOVA: `svyglm` Syntax

- As with t-tests, regressions are done in the package `survey` not `srvyr` but you can use the same design object

- Syntax is similar between t-test and glm

```{r glm_syn, eval=FALSE}
svyglm(formula, 
       design,
       na.action, #default is na.omit
       ....)
```
---
## `svyglm` Example: Two-sample

Same example as two-sample t-test: Are electric bills different between those with and without A/C?

<b>t-test:</b>
```{r twosamp_ex_ttest, eval=FALSE}
recs_des %>%
   svyttest(design=.,
            formula=DOLLAREL~ACUsed,
            na.rm=TRUE) #<<
```

<b>glm:</b>
```{r twosamp_ex_glm, eval=FALSE}
recs_des %>%
   svyglm(design=.,
          formula=DOLLAREL~ACUsed,
          na.action=na.omit) #<<
```

???
- one major difference in how you specify to ignore NA values
- svyttest can only have 2-levels in group variable
- svyglm, the variable on right can be anything (continuous or factor)

---
## `svyglm` Example: Two-sample

Are electric bills different between those with and without A/C?
.small[
```{r twosamp_ex_ttest_run}
recs_des %>%
   svyglm(design=.,
          formula=DOLLAREL~ACUsed,
          na.action=na.omit) %>%
  summary()
```
]

???
- same results as t-test

---
## `svyglm` Example 1: ANOVA Test

Does temperature of AC at night vary by region?
.smaller[
```{r anova_ex}
recs_des %>%
   svyglm(design=.,
          formula=SummerTempNight~Region,
          na.action=na.omit) %>%
  summary()

```
]

???
- Region is a factor variable, if it is numeric - this will treat it like a linear model

---
## `svyglm` Example 2: Linear Model

- Is there a relationship between square footage and electric bill?
- Let's review the data first with a ggplot. <i>Note we use the original data and do <b>NOT</b> use the survey design object.</i>

```{r plot_sf_elbill}
p <- recs %>%
  ggplot(aes(x=TOTSQFT_EN, y=DOLLAREL, weight=NWEIGHT)) +
  geom_hex() + 
  theme(legend.position="right") +
  guides(fill=guide_legend(title="HUs"))
```
---
## `svyglm` Example 2: Linear Model
```{r plot_sf_elbill_disp, echo=FALSE, fig.asp=9/16, fig.align="center", out.width="90%", dpi=300}
p +
   theme_xaringan() 
```


---
## `svyglm` Example 2: Linear Model
.small[
```{r lm_ex}
m_electric_sqft <- recs_des %>%
   svyglm(design=.,
          formula=DOLLAREL~TOTSQFT_EN,
          na.action=na.omit)
summary(m_electric_sqft)
```
]

???
- for every square foot bigger, on average 24.6c more in electric
---
## Practice on your own

- Open ContinuousExercises.Rmd and work through Part 2

- We will take 15 minutes. Use this time for the exercises and a break
---
class: inverse center middle
# Categorical data analysis
---
## Weighted Analysis for Categorical Variable

- Functions to use within `summarize` after `group_by`
   - survey_mean
   - survey_total

- Functions to get counts
  - survey_count

???

- we use the same mean and total functions as with continuous variables
- `survey_count` is new
    - has a similar structure as the standard (non-survey) version of count

---
## Set-up ANES Data for Examples

```{r anes_des}
anes <- read_rds(here("Data", "anes.rds")) %>%
   mutate(Weight=Weight/sum(Weight)*224059005) 
# adjust weight to sum to citizen pop, 18+ in Nov 2016 per ANES methodology documentation
anes_des <- anes %>%
   as_survey_design(weights = Weight,
                    strata = Stratum,
                    ids = VarUnit,
                    nest = TRUE)

```
???

- American National Election Studies
- provides weights that sum to the sample, but we want to get population estimates
    - need to adjust the weight to get it to the population count
- as we mentioned before we will cover setting up the sample design object later

---
## `survey_count` Syntax

- `survey_count` functions similarly to `count` in that it is <b>NOT</b> called within `summarize`

- Produces weighted counts and variance of your choice of those counts
```{r survey_count_syn, eval=FALSE}
survey_count(
  x,
  ...,
  wt = NULL,
  sort = FALSE,
  name = "n",
  .drop = dplyr::group_by_drop_default(x),
  vartype = c("se", "ci", "var", "cv")
)
```
???
- similar to count in that it takes one or many variables
- can change the variance type as we have seen in the other survey functions

---
## `survey_count` Example

- Cross-tab of population in each age group and gender
```{r survey_count_ex}
anes_des %>%
  survey_count(AgeGroup, Gender, name="n")

```
???
- `survey_count` is placed on its own like `count`
- it does NOT go in a `summarize` function
- can take multiple variables
- can change the output count name, `n` is the default

---
## `survey_mean` and `survey_total` Examples

- `survey_mean` used with no x (variable) calculates a proportion of groups specified in `group_by`
- `survey_total` used with no x (variable) calculates a population count estimate for the groups specified in `group_by`

Cross-tab of population who voted in 2016
```{r survey_p_ex1}
anes_des %>%
  filter(!is.na(VotedPres2016)) %>%
  group_by(VotedPres2016) %>%
  summarize(
    p=survey_mean(),
    N=survey_total(),
    n=unweighted(n()), .groups="drop"
  )
```
???
- to get proportions we use `group_by` and `survey_mean`
- also use `survey_total` to get a population count estimate as before
---
## Conditional proportions with more than one group

- Specifying more than one group calculates conditional proportions
- Example: people voting in 2012 and 2016

```{r survey_p_cond}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(VotedPres2012, VotedPres2016) %>%
  summarize(
    p=survey_mean(),
    N=survey_total(),
    n=unweighted(n()), .groups="drop"
  )
```
???
- Note that this is the proportion of voting in 2016 by whether people voted in 2012
- What if we don't want conditional proportions?

---
## Joint proportions with more than one group

- Specify an interaction to get joint distribution
- Example: people voting in 2012 and 2016

```{r survey_p_joint}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(groups = interaction(VotedPres2016, VotedPres2012)) %>% #<<
  summarize(
    p=survey_mean(),
    N=survey_total(),
    .groups="drop"
  )
```
???
- We add an interaction for the groups
- This outputs the joint distribution, but the `groups` variable is hard to interpret

---
## Joint proportions with more than one group

- Specify an interaction to get joint distribution
- Example: people voting in 2012 and 2016

```{r survey_p_joint2}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(groups = interaction(VotedPres2016, VotedPres2012)) %>% 
  summarize(
    VotedPres2012=VotedPres2012[1], #<<
    VotedPres2016=VotedPres2016[1], #<<
    p=survey_mean(),
    N=survey_total(),
    .groups="drop"
  )
```
???
- We can add in two variables one for `VotedPres2012` and `VotedPres2016`
- using the 1 in brackets pulls out the labels for these two variables so we see the "Yes" and "No" labels

---
## Proportions with Design Effects

```{r survey_p_deff}
anes_des %>%
  filter(!is.na(VotedPres2012), !is.na(VotedPres2016)) %>%
  group_by(VotedPres2012, VotedPres2016) %>%
  summarize(
    p=survey_mean(deff=TRUE),
    N=survey_total()
  )
```
???
- Use `deff=TRUE` option in the `survey_mean` function

---
## `svychisq` Syntax

- As with testing on continuous variables, `svychisq` comes from the `survey` package

```{r svychisq_syn, eval=FALSE}
svychisq(formula,
         design, 
         statistic = c("F",  "Chisq", "Wald", "adjWald", "lincom", "saddlepoint"),
         na.rm=TRUE,
         ...)
                       
```
???
- when we want to test categorical distributions we use `svychisq`
- it takes a formula, and the survey design data

---
## `svychisq` Example 1: Function Defaults

- How often can you trust the federal gov't to do what is right?
- How often can you trust other people?

```{r svychisq_ex1}
anes_des %>%
   svychisq(design=.,
            formula=~TrustPeople +TrustGovernment)


```
???
- We want to compare the distributions of these two questions

---
## `svychisq` Example 2: Wald Statistic

- How often can you trust the federal gov't to do what is right?
- Who did you vote for? Clinton, Trump, or Other

```{r svychisq_ex2}
anes_des %>%
   svychisq(design=.,
            formula=~TrustGovernment +VotedPres2016_selection,
            statistic="Wald")


```
???
- Can use different statistics
---
## Practice on your own

- Open CategoricalExercises.Rmd and work through the exercises

- We will take 10 minutes. Use this time for the exercises and a break
---
class: inverse center middle
# Sample design object
---
## `tbl_svy` Object: Taylor's Series

- `tbl_svy` object defines the sampling design or replicate weights

- Key information is usually found in documentation of a public use file

```{r sd_tsl_syn, eval=FALSE}
as_survey_design(
  .data,
  ids = NULL,#cluster IDs/PSUs
  strata = NULL,#strata variables
  variables = NULL,#defaults to all in .data
  fpc = NULL,#variables defining the fpc
  nest = FALSE,#TRUE/FALSE - relabel clusters to nest within strata
  check_strata = !nest, #check that clusters are nested in strata
  weights = NULL,# weight variable
  ...
)
```

???
- discussing TSL first
---
## `tbl_svy` for Common Designs

```{r sd_tsl_gen_ex, eval=FALSE}
# simple random sample (SRS)
apisrs %>% as_survey_design(fpc = fpc)

# stratified sample
apistrat %>% as_survey_design(strata = stype, weights = pw)

# one-stage cluster sample
apiclus1 %>% as_survey_design(ids = dnum, weights = pw, fpc = fpc)

# two-stage cluster sample, weights computed from pop size
apiclus2 %>% as_survey_design(ids = c(dnum, snum), fpc = c(fpc1, fpc2))

# stratified, cluster design
apistrat %>% as_survey_design(ids = dnum, strata = stype, weights =pw, nest = TRUE)

```

- examples from `srvyr` help documentation

---
## ANES Design Object

.smaller[
```{r sd_anes, eval=TRUE}
anes_des <- anes %>%
   as_survey_design(weights = Weight,
                    strata = Stratum,
                    ids = VarUnit,
                    nest = TRUE)
summary(anes_des)
```
]

???
- Pseudo-strata (Stratum) and pseudo-cluster (VarUnit) included for variance estimation
- we renamed these variables to be more obvious but original documentation has diff var names

---
## `tbl_svy` Objects with Supplied Replicate Weights

- Key information is usually found in documentation of a public use file

```{r sd_rep_syn, eval=FALSE}
as_survey_rep(
  .data,
  variables = NULL,#defaults to all in .data
  repweights = NULL,#variables specifying replicate weights
  weights = NULL,#variable for analysis weight
  type = c("BRR", "Fay", "JK1", "JKn", "bootstrap", "other"),
  rho = NULL,#shrinkage factor for Fay's method,
  mse = getOption("survey.replicates.mse"), #	if TRUE, compute variances based on 
  # sum of squares around the point estimate, rather than the mean of the replicates
  scale = NULL, # overall multiplier for squared deviations
  ...
)

```
---
## RECS Design Object

.smaller[
```{r sd_recs, eval=TRUE}
recs_des <- recs %>%
   as_survey_rep(weights=NWEIGHT,
                 repweights=starts_with("BRRWT"),
                 type="Fay",
                 rho=0.5,
                 mse=TRUE)
summary(recs_des)
```
]

???
- Fay's method of BRR weight with $\epsilon=0.5$
- RECS documentation includes syntax for creating survey design object
- https://www.eia.gov/consumption/residential/data/2015/pdf/microdata_v3.pdf

---
## Create Replicate Weights: jackknife

- You can also start with a design object specified by the design and create replicate weights
.smaller[
```{r sd_create_rep}
data(api)
dclus1 <- apiclus1 %>% as_survey_design(ids = dnum, weights = pw, fpc = fpc)
rclus1 <- as_survey_rep(dclus1)
summary(rclus1)

```
]
---
## Create Replicate Weights: bootstrap

- You can also start with a design object specified by the design and create replicate weights
.small[
```{r sd_create_boot}
bclus1 <- as_survey_rep(dclus1, type="bootstrap", replicates=100)
summary(bclus1)

```
]
---
## Create Survey Design Object for ACS

Fill in the blanks
- Analysis weight: PWGTP
- replicate weights: PWGTP1-PWGTP180
- jackknife with scale adjustment of 4/80
```{r sd_acs_fib, eval=FALSE}
acs_des <- acs_pums %>%
  as_survey_rep(
    weights=___________,
    repweights=___________,
    type=___________,
    scale=_________
  )
```
--

```{r sd_acs_fib_sol, eval=FALSE}
acs_des <- acs_pums %>%
   as_survey_rep(
      weights=PWGTP,
      repweights=stringr::str_c("PWGTP", 1:80),
      type="JK1",
      scale=4/80
   )

```
---
## Create Survey Design Object for CPS 2011 Supplement

Fill in the blanks
- Analysis weight: wtsupp
- replicate weights: repwtp1 -repwtp160
- BRR
```{r sd_cps_fib, eval=FALSE}
cps_des <- cps %>%
  as_survey_rep(
    weights=___________,
    repweights=___________,
    type=___________
  )
```
--
```{r sd_cps_fib_sol, eval=FALSE}
cps_des <- cps %>%
  as_survey_rep(
    weights=wtsupp,
    repweights=starts_with("repwtp"),
    type="BRR"
  )
```
---
## Create Survey Design Object for NHANES

Fill in the blanks
- Analysis weight: WTINT2YR
- Variance Stratum: SDMVSTRA
- Variance Primary Sampling Unit: VPSU
```{r sd_nhanes_fib, eval=FALSE}
nhanes_des <- nhanes %>%
  as_survey_design(
    weights=___________,
    ids=___________,
    strata=___________,
    fpc=___________
  )
```
--
```{r sd_nhanes_fib_sol, eval=FALSE}
nhanes_des <- nhanes %>%
  as_survey_design(
    weights=WTINT2YR,
    ids=VPSU,
    strata=SDMVSTRA,
    fpc=NULL
  )
```
---
## Create Survey Design Object for LEMAS 2016

Fill in the blanks
- Analysis weight: ANALYSISWEIGHT
- Variance Stratum: STRATA
- FPC: FRAMESIZE
```{r sd_lemas_fib, eval=FALSE}
lemas_des <- lemas %>%
  as_survey_design(
    weights=___________,
    ids=___________,
    strata=___________,
    fpc=___________
  )
```
--

```{r sd_lemas_fib_sol, eval=FALSE}
lemas_des <- lemas %>%
  as_survey_design(
    weights=ANALYSISWEIGHT,
    ids=1,
    strata=STRATA,
    fpc=FRAMESIZE
  )
```

---
class: inverse center middle
# Closing
---
## Resources for more learning

- https://cran.r-project.org/web/packages/srvyr/vignettes/srvyr-vs-survey.html

- https://r-survey.r-forge.r-project.org/survey/ 
  - Includes more advanced modeling


---
## Thank You!

### We hope you learned a lot in this short course!

Please let us know if you have any feedback on this course.  You will receive an email from AAPOR asking you to fill out a survey about this course. All feedback is welcome!


## Questions?

---
## Sources

- <font size="2">The American National Election Studies (https://electionstudies.org/). These materials are based on work supported by the National Science Foundation under grant numbers SES 1444721, 2014-2017, the University of Michigan, and Stanford University.  </font>

- <font size="2">*Residential Energy Consumption Survey: Using the 2015 Microdata File to Compute Estimates and Standard Errors.* U.S. Department of Energy (2017) https://www.eia.gov/consumption/residential/data/2015/pdf/microdata_v3.pdf </font>

- <font size="2">Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0. https://allisonhorst.github.io/palmerpenguins/ </font>

- <font size="2">T. Lumley (2020) "survey: analysis of complex survey samples". R package version 4.0. https://r-survey.r-forge.r-project.org/survey/ </font>

- <font size="2">Greg Freedman Ellis and Ben Schneider (2020). srvyr: 'dplyr'-Like Syntax for Summary Statistics of Survey Data. R package version 1.0.0. https://CRAN.R-project.org/package=srvyr </font>

- <font size="2">Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2021). dplyr: A Grammar of Data Manipulation. R package version 1.0.5. https://CRAN.R-project.org/package=dplyr </font>
