---
title: "Categorical Data Analysis Exercise Solutions"
output:
  html_document:
    df_print: paged
---

# Set-up
```{r}
library(tidyverse) # for tidyverse
library(here) # for file paths
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

anes <- read_rds(here("Data", "anes.rds")) %>%
   mutate(Weight=Weight/sum(Weight)*224059005) 
# adjust weight to sum to citizen pop, 18+ in Nov 2016 per ANES methodology documentation

anes_des <- anes %>%
   as_survey_design(weights = Weight,
                    strata = Stratum,
                    ids = VarUnit,
                    nest = TRUE)
```

# Part 1

1. How many females have a graduate degree?

```{r}


```

2. What percentage of people identify as "Strong democrat"?

```{r}

```

3. What percentage of people who voted in the 2016 election identify as "Strong republican"?

```{r}

```

4. What percentage of people voted in both the 2012 election and in the 2016 election?  Include the confidence interval.

```{r}

```

5. What is the design effect for the proportion of people who voted early?

```{r}

```

# Part 2

1. Is there a relationship between PartyID and When people voted in the 2016 election (on election day or early voting)?

```{r}

```

2. Is there a relationship between PartyID and trust in the government?

```{r}

```


# Bonus

1. What percentage of people lean republican?

```{r}

```

2. Were people who lean democrat more likely to vote early in the 2020 election?

```{r}

```