---
title: "Categorical Data Analysis Exercise Solutions"
output:
  html_document:
    df_print: paged
---

# Set-up
```{r}
library(tidyverse) # for tidyverse
library(here) # for file paths
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

anes <- read_rds(here("Data", "anes.rds")) %>%
   mutate(Weight=Weight/sum(Weight)*224059005) 
# adjust weight to sum to citizen pop, 18+ in Nov 2016 per ANES methodology documentation

anes_des <- anes %>%
   as_survey_design(weights = Weight,
                    strata = Stratum,
                    ids = VarUnit,
                    nest = TRUE)
```

# Part 1

1. How many females have a graduate degree?

```{r}
#Option 1:
anes_des %>%
  filter(Gender=="Female", Education=="Graduate") %>%
  survey_count(name="n")
#Option 2:
anes_des %>%
  filter(Gender=="Female", Education=="Graduate") %>%
  summarize(
    N=survey_total(),  .groups="drop"
  )

```

2. What percentage of people identify as "Strong democrat"?

```{r}
anes_des %>%
  group_by(PartyID) %>% 
  summarize(
    p=survey_mean()
  ) %>%
  filter(PartyID=="Strong democrat")
```

3. What percentage of people who voted in the 2016 election identify as "Strong republican"?

```{r}
anes_des %>%
  filter(VotedPres2016=="Yes") %>% 
  group_by(PartyID) %>% 
  summarize(
    p=survey_mean()
  )
```

4. What percentage of people voted in both the 2012 election and in the 2016 election?  Include the confidence interval.

```{r}
anes_des %>%
  group_by(groups=interaction(VotedPres2012, VotedPres2016)) %>% 
  filter(!is.na(groups)) %>%
  summarize(
    VotedPres2012=VotedPres2012[1],
    VotedPres2016=VotedPres2016[1],
    p=survey_mean(var="ci")
  )
```

5. What is the design effect for the proportion of people who voted early?

```{r}
anes_des %>%
   filter(!is.na(EarlyVote2016)) %>%
   group_by(EarlyVote2016) %>%
   summarize(
      p=survey_mean(deff=TRUE),
      N=survey_total()
   )
```

# Part 2

1. Is there a relationship between PartyID and When people voted in the 2016 election (on election day or early voting)?

```{r}
anes_des %>%
   svychisq(design=.,
            formula=~PartyID +EarlyVote2016)
```

2. Is there a relationship between PartyID and trust in the government?

```{r}
anes_des %>%
   svychisq(design=.,
            formula=~PartyID+TrustGovernment,
            statistic="Wald")
```


# Bonus

1. What percentage of people lean republican?

```{r}

#Solution 1: Using forcats package
anes_des %>%
   mutate(PartyID3=fct_collapse(PartyID,
                                LeanDem=c("Strong democrat",
                                          "Not very strong democrat",
                                          "Independent-democrat"),
                                LeanRep=c("Strong republican",
                                          "Not very strong republican",
                                          "Independent-republican"),
                                other_level="Other")) %>% 
   group_by(PartyID3) %>% 
   summarize(p=survey_mean())

#Solution 2: Using case_when
anes_des %>%
   mutate(PartyID3=case_when(PartyID %in% c("Strong democrat",
                                            "Not very strong democrat",
                                            "Independent-democrat")~"LeanDem",
                             PartyID %in% c("Strong republican",
                                            "Not very strong republican",
                                            "Independent-republican")~"LeanRep",
                             is.na(PartyID)~NA_character_,
                             TRUE~"Other")) %>% 
   group_by(PartyID3) %>% 
   summarize(p=survey_mean())

```

2. Were people who lean democrat more likely to vote early in the 2020 election?

```{r}

earlyv_glm<-anes_des %>%
   mutate(PartyID3=fct_collapse(PartyID,
                                LeanDem=c("Strong democrat",
                                          "Not very strong democrat",
                                          "Independent-democrat"),
                                LeanRep=c("Strong republican",
                                          "Not very strong republican",
                                          "Independent-republican"),
                                other_level="Other")) %>% 
   svyglm(design=.,
          formula=(EarlyVote2016=="Yes")~PartyID3,
          family=quasibinomial(),
          na.action=na.omit)

summary(earlyv_glm)
```