---
title: "Continous Data Analysis Exercises"
output:
  html_document:
    df_print: paged
---

# Set-up
```{r}
library(tidyverse) # for tidyverse
library(here) # for file paths
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

recs <- read_rds(here("Data", "recs.rds"))

recs_des <- recs %>%
   as_survey_rep(weights=NWEIGHT,
                 repweights=starts_with("BRRWT"),
                 type="Fay",
                 rho=0.5,
                 mse=TRUE)
```

# Part 1

1. Find the average square footage of housing units (TOTSQFT_EN) with a 90% confidence interval.

```{r}

```

2. Estimate the ratio of cooled square footage to total square footage (TOTCSQFT) to the total square footage of housing units (TOTSQFT_EN) with its standard error.

```{r}

```

3. Estimate the median temperature housing units are set to during the night in the winter (WinterTempNight) using the `survey_median` function.

```{r}

```

4. Estimate the median temperature housing units are set to during the night in the winter (WinterTempNight) using the `survey_quantile` function.

```{r}

```

# Part 2

1. Estimate the total average energy cost (TOTALDOL) by region, division, and urbanicity.

```{r}

```

2. What is the median electric cost (DOLLAREL) for housing units in the South Region? What is the 95% confidence interval?

```{r}

```

3. Test whether daytime winter and daytime summer temperatures of homes are set the same.

```{r}

```

4. Test whether average electric bill (DOLLAREL) varies by region (Region).

```{r}

```

5. Fit a regression between the cooled square footage of a housing unit (TOTCSQFT) and the total amount spent on energy (TOTALDOL).

```{r}

```

