---
title: "Warm-up Exercise Solutions"
output:
  html_document:
    df_print: paged
---

# Course set-up
First, let's make sure you have everything you need for the course. Run the following library statements. If something is not installed, install it.

```{r}
# install.packages("tidyverse")
# install.packages("srvyr")
# install.packages("here")
# install.packages("palmerpenguins")
# install.packages("remotes")

library(tidyverse) # for tidyverse
library(here) # for file paths

remotes::install_github("bschneidr/survey", ref = "c217689")
library(srvyr)
library(palmerpenguins)


```

# Warm-up exercises: Play with penguin data!!!

```{r}
penguins
```

How many penguins of each species are there? Hint: use `count`
```{r}
penguins %>%
   count(species)
```

How many penguins of each species and sex are there? Hint: use `count`

```{r}
penguins %>%
   count(species, sex)
```

What is the mean length of flipper by species? Hint: use `group_by` and `summarise`
```{r}
penguins %>%
   group_by(species) %>%
   summarize(
     MeanFlipperLength=mean(flipper_length_mm, 
                            na.rm=TRUE))
```


What is the mean flipper length by species and sex? Hint: use `group_by` and `summarise`
```{r}
penguins %>%
   group_by(species, sex) %>%
   summarize(
     MeanFlipperLength=mean(flipper_length_mm,
                            na.rm=TRUE))
```

# Advanced warm-up exercises

Fit a simple linear regression between body mass and flipper length.

```{r}
mod1 <- lm(body_mass_g ~ flipper_length_mm,
           data=penguins)
summary(mod1)
```


Test whether the average flipper length is significantly different between male and female penguins. Use t-test, lm, or glm

```{r}
t.test(flipper_length_mm ~ sex, data=penguins)

mod3 <- lm(flipper_length_mm ~ sex, data=penguins)
summary(mod3)

mod4 <- glm(flipper_length_mm ~ sex, data=penguins)
summary(mod4)
```

